Algebraic File Synchronization: Adequacy and Completeness

### 2017-10-19 p2/axioms.tex
Completed the completeness theorem
Added TODO comments
Various improvements

So that we can describe the effects of commands independently of filesystems,
let us introduce some notation
and note some observations.
We already know that
commands usually do not occur in isolation,
and are applied to filesystems in time.
Therefore we investigate sequences of commands with a well-defined order.
\begin{mydef}[Sequences of commands: $\cc$ and $\emptyseq$]
We use ``$\cc$'' to combine commands to form a sequence, or combine sequences to form a longer sequence,
with the meaning that the command on the left is applied first:
\[ \alpha\cc\beta\aFS = \beta(\alpha\aFS). \]
%% monoid, but not free monoid
%% monoid: semigroup with identity element
%% Sequences of commands form a free semigroup
Sequences are also partial endofunctions on filesystems, defined only if all commands it contains succeed in the given order.
We write $\sets$ for the set of sequences of commands.
They form a monoid, and, as usual,
$\emptyseq$ denotes its unit element, the empty sequence,
which is defined on all filesystems and by definition leaves all filesystems unchanged.
\end{mydef}

@@

\begin{myax}\mlabel{ax_directchild_breaks}
All other commands on a child break every filesystem:
$\cxynnv\cc\czwnv \equiv \cbrk$ where $\nn=\parent(n)$ and $\cxy\neq\cdd$ and $\czw\neq\cbb$
and the pair is not a construction pair.
As we shall see, this can be extended to \cref{simple_distant_pairs}.
% TODO Delete this rule
\end{myax}

@@

\begin{myax}\mlabel{ax_directparent_breaks}
All other commands on a parent break every filesystem:
$\cxynv\cc\czwnnv \equiv \cbrk$ where $\parent(n)=\nn$ and $\cxy\neq\cbb$ and $\czw\neq\cdd$
and the pair is not a destruction pair.
This can also be extended to \cref{simple_distant_pairs}.
% TODO Delete this rule
\end{myax}

@@

If $n=m$, then from \cref{ax_same_breaks} we know $[y]=[z]$ and $[w]=[x]$
as otherwise the commands, in one order or the other, would break all filesystems
and they could not be independent.
The commands are not assertion commands, so from \cref{ax_same_singlec}
we know that $\cxynv\cc\czwnv\equiv\czwnv\cc\cxynv$ can only happen if $y=w$.
% TODO Does this depend on |Vd|=1?
Therefore $[x]=[y]=[z]=[w]$ which with $y=w$ contradicts our assumption that the commands are different.

>>>>

So that we can describe the effects of commands independently of filesystems,
let us introduce some notation
and note some observations.
We already know that
commands usually do not occur in isolation,
and are applied to filesystems in time.
Therefore we investigate sequences of commands with a well-defined order.
\begin{mydef}[Sequences of commands: $\cc$ and $\emptyseq$]
We use ``$\cc$'' to combine commands to form a sequence, or combine sequences to form a longer sequence,
with the meaning that the command on the left is applied first:
\[ \alpha\cc\beta\aFS = \beta(\alpha\aFS). \]
%% monoid, but not free monoid
%% monoid: semigroup with identity element
%% Sequences of commands form a free semigroup
Sequences are also partial endofunctions on filesystems,
defined only if all commands they contain succeed in the given order.
We write $\sets$ for the set of sequences of commands.
They form a monoid, and, as usual,
$\emptyseq$ denotes its unit element, the empty sequence,
which is defined on all filesystems and by definition leaves all filesystems unchanged.
\end{mydef}

@@

\begin{myax}\mlabel{ax_directchild_breaks}
All other commands on a child break every filesystem:
$\cxynnv\cc\czwnv \equiv \cbrk$ where $\nn=\parent(n)$ and $\cxy\neq\cdd$ and $\czw\neq\cbb$
and the pair is not a construction pair.
As we shall see, this can be extended to \cref{simple_distant_pairs}.
TODO TO BE DELETED -- WE ONLY WANT TO USE \cref{simple_distant_pairs}
\end{myax}

@@

\begin{myax}\mlabel{ax_directparent_breaks}
All other commands on a parent break every filesystem:
$\cxynv\cc\czwnnv \equiv \cbrk$ where $\parent(n)=\nn$ and $\cxy\neq\cbb$ and $\czw\neq\cdd$
and the pair is not a destruction pair.
This can also be extended to \cref{simple_distant_pairs}.
TODO TO BE DELETED -- WE ONLY WANT TO USE \cref{simple_distant_pairs}
\end{myax}

@@

If $n=m$, then from \cref{ax_same_breaks} we know $[y]=[z]$ and $[w]=[x]$
as otherwise the commands, in one order or the other, would break all filesystems
and they could not be independent.
The commands are not assertion commands, so from \cref{ax_same_singlec}
we know that $\cxynv\cc\czwnv\equiv\czwnv\cc\cxynv$ can only happen if $y=w$.
% TODO Does this depend on |[D]|=1?
Therefore $[x]=[y]=[z]=[w]$ which with $y=w$ contradicts our assumption that the commands are different.

### 2017-10-19 p2/ordering.tex
Various improvements

\begin{mydef}[$\ordersetsign$]\mlabel{def_orderset}
For a simple set or sequence of commands $S$,
we use $\orderset{S}$ to denote the set of sequences
that are all permutations of the commands in $S$
for which the following holds:
for all subsequences $\cxynv\cc\czwmv$,
\begin{itemize}
\item if $n=\parent(m)$, then $y=\vald$
\item if $\parent(n)=m$, then $y=\empt$.
\end{itemize}
\end{mydef}

@@

The principal ideas behind these conditions are 
\cref{ax_separate_commute}, from which we know that commands on incomparable
nodes can be applied in any order, so we only need to focus on the order of commands
with comparable nodes;
\cref{connected_changes}, from which we know that if $U$ contains commands on two comparable
nodes, then it contains commands on all nodes in between, so it is enough to specify
the order of commands on directly related nodes;
and finally, \cref{simple_distant_pairs}, from which we know
that commands on parent--child pairs must be construction or destruction pairs.
Investigating these pairs we see that they are identified by the output value of the first command ($y$),
which in turn determines whether the command on the parent or the child must precede the other.

@@

\begin{mycor}\mlabel{order_is_only_possible}
For any simple sequence $S$ and its permutation $S'$, 
\[ S'\not\in\orderset{S} \Longrightarrow S'\equiv\cbrk. \]
\end{mycor}
\begin{proof}
This follows directly from \cref{simple_distant_pairs}.
If $S'\not\in\orderset{S}$ then by \cref{def_orderset}
it contains two commands on a node $n$ and its parent
that placed next to each other do not form
a construction or destruction pair.
Let them be $\cxynv$ and $\czwmv$.
By \cref{ax_directchild_breaks,ax_directparent_breaks}
we know
$\cxynv\cc\czwmv\equiv\cbrk$, which is a subsequence of $S'$,
and therefore by \cref{simple_distant_pairs}, $S'\equiv\cbrk$.
\end{proof}

@@

\begin{mylem}\mlabel{simple_reorder_equiv}
For a simple sequence $\wrksx{S}$,
$\orderset{S}$ is the set of all simple sequences equivalent to $S$, that is,
\[ \orderset{S} = [S]_\equiv \cap \setssmp. \]
\end{mylem}

@@

% TODO APPENDIX
\begin{proof}
First, we prove, by contradiction, that if a simple sequence $T\equiv S$, then $T\in\orderset{S}$.
We assume $T\equiv S$ and $T\not\in\orderset{S}$.
Then, from \cref{equiv_simple_same_commands} we know $T$ is a permutation of $S$,
and from \cref{order_is_only_possible} we know $T\equiv\cbrk$ which
is a contradiction as $T\equiv \wrksx{S}$.

@@

In our nested induction step we assume $T'\equiv S$ for all $T'\in\orderset{S}$
where there are at most $j$ commands in $T'$ to the left of $\cxynv$.
Let us then consider $T$ where there are $j+1$ such commands.
We aim to transform $T$ into $T''\equiv T$ by swapping $\cxynv$ with the preceding
command. Then from the induction hypothesis we will know $T\equiv T''\equiv S$,
which proves our lemma.

@@

It can be shown in the same way
that the case when the $n_\indx$ nodes ascend the tree,
that is, $n_{\indx+1}=\parentf{n_\indx}$ is true, also leads to a contradiction,
by using destruction commands instead of construction ones.
\end{proof}

\begin{mydef}[$\ordsign$]
It follows from \cref{simple_reorder_equiv} that given a simple set of commands $S$,
which can be applied to a filesystem in some order without breaking it,
this order will be in $\orderset{S}$,
and all sequences in $\orderset{S}$
represent the same partial endofunction on filesystems.
This allows us to treat such a set as a sequence for any purpose
where the internal order of the commands is irrelevant.
We write $\ordp{S}$ to represent this endofunction.
Also, let $\ordp{S}=\cbrk$ if such an order does not exist.
\end{mydef}

>>>>

\begin{mydef}[$\ordersetsign$]\mlabel{def_orderset}
For a simple set or sequence of commands $S$,
we use $\orderset{S}$ to denote the set of sequences
that are all permutations of the commands in $S$
for which the following holds:
for all subsequences $\cxynv\cc\czwmv$,
\begin{itemize}
\item if $n=\parent(m)$, then $[y]=[\vald]$
\item if $\parent(n)=m$, then $y=\empt$.
\end{itemize}
\end{mydef}

@@

The principal ideas behind these conditions are 
\cref{ax_separate_commute}, from which we know that commands on incomparable
nodes can be applied in any order, so we only need to focus on the order of commands
with comparable nodes;
\cref{connected_changes}, from which we know that if a simple sequence contains commands on two comparable
nodes, then it contains commands on all nodes in between, so it is enough to specify
the order of commands on directly related nodes;
and finally, \cref{simple_distant_pairs}, from which we know
that commands on parent--child pairs must be construction or destruction pairs.
Investigating these pairs we see that they are identified by the output value of the first command ($y$),
which in turn determines whether the command on the parent or the child must precede the other.

@@

\begin{mycor}\mlabel{order_is_only_possible}
For any simple sequence $S$ and its permutation $S'$, 
\[ S'\not\in\orderset{S} \Longrightarrow S'\equiv\cbrk. \]
\end{mycor}
\begin{proof}
This follows directly from \cref{simple_distant_pairs}.
We use an inverse proof and assume $\wrks{S'}$.
If $S'\not\in\orderset{S}$ then by \cref{def_orderset}
it contains two commands on a node $n$ and its parent
that, as a subsequence,
do not form a construction or destruction pair.
However, this contradicts \cref{simple_distant_pairs}
which means that no such subsequence exists.
\end{proof}

@@

\begin{mylem}\mlabel{simple_reorder_equiv}
For a simple sequence $\wrksx{S}$,
$\orderset{S}$ is the set of all simple sequences equivalent to $S$, that is,
\[ \orderset{S} = [S]_\equiv \cap \setssmp, \]
where $\setssmp$ is the set of all simple sequences.
\end{mylem}

@@

% TODO APPENDIX
\begin{proof}
First, we prove, by contradiction, that if a simple sequence $T$, $T\equiv S$, then $T\in\orderset{S}$.
We use an inverse proof and assume $T\equiv S$ and $T\not\in\orderset{S}$.
Then, from \cref{equiv_simple_same_commands} we know $T$ is a permutation of $S$,
and from \cref{order_is_only_possible} we know $T\equiv\cbrk$ which
is a contradiction as $T\equiv \wrksx{S}$.

@@

In our nested induction step we assume $T'\equiv S$ for all $T'\in\orderset{S}$
where there are at most $j$ commands in $T'$ to the left of $\cxynv$.
Let us then consider $T$ where there are $j+1$ such commands.
We aim to transform $T$ into $T''\equiv T$ by swapping $\cxynv$ with the preceding
command. Then from the induction hypothesis we know $T\equiv T''\equiv S$,
which proves our lemma.

@@

It can be shown in the same way
that the case when the $n_\indx$ nodes ascend the tree,
that is, $n_{\indx+1}=\parentf{n_\indx}$ is true, also leads to a contradiction.
In this case we find that all commands are destruction commands,
and so $[y_\indx]\neq[\vald]$, while in a subsequence of $T$ for a given $\indx$,
$[y_\indx]=[\vald]$ must be true by the definition of $\ordersetsign$.
\end{proof}

\begin{mydef}[$\ordsign$]
It follows from \cref{simple_reorder_equiv} that given a simple set of commands $S$,
which can be applied to a filesystem in some order without breaking it,
this order will be in $\orderset{S}$,
and all sequences in $\orderset{S}$
represent the same partial endofunction on filesystems.
This allows us to treat such a set as a sequence for any purpose
where the internal order of the commands is irrelevant.
We write $\ordp{S}$ to represent this endofunction.
We leave $\ordp{S}$ undefined if such an order does not exist.
\end{mydef}

### 2017-10-19 p2/upd-correct.tex
Various improvements

\begin{myth}\mlabel{update_works}
A sequence of commands returned by the update detector
when comparing the non-broken $\FS^*$ to the original $\FS$
will not break $\FS$ if it is applied to it.
\end{myth}
\begin{proof}
Let $\FS\neq\fsbroken$ be the original filesystem, and $\FS^*\neq\fsbroken$
be the filesystem after the changes we intend to detect.
Let us assume that some mechanism recorded all changes that occurred
to $\FS$ until it reached the state $\FS^*$.
Our set if commands is sufficient to record any change, as
there are commands for every input and output type pair, therefore
we know that the changes can be recorded as a sequence of commands $S$
where $S\aFS=\FS^*$.
Note that $S$ will not contain assertion commands, as they do not
represent an actual change in the filesystem.

@@

>>>>

% TODO
TODO ENSURE UPDATE DETECTOR IS DEFINED.

\begin{myth}\mlabel{update_works}
A sequence of commands returned by the update detector
when comparing the non-broken $\FS^*$ to the original $\FS$
will not break $\FS$ if it is applied to it.
\end{myth}
\begin{proof}
Let $\FS$ be the original filesystem, and $\FS^*$
be the filesystem after the changes we intend to detect.
Let us assume that some mechanism recorded all changes that occurred
to $\FS$ until it reached the state $\FS^*$.
Our set if commands is sufficient to record any change, as
there are commands for every input and output type pair, therefore
we know that the changes can be recorded as a sequence of commands $S$
where $S\aFS=\FS^*$.
Note that $S$ will not contain assertion commands, as they do not
represent an actual change in the filesystem.

@@

TODO TODO TODO

### 2017-10-19 p2/upd-def.tex
Rename "works" to "defined"
Various improvements

\begin{mydef}[Minimal sets or sequences, $\setsmin$]
A sequence or set of commands is minimal if it contains at most one command on each node.
We write $\setsmin$ for the set of minimal sequences.
\end{mydef}

\begin{mydef}[Simple sets or sequences, $\setssmp$]
A sequence or set of commands is simple if it is minimal and it does not contain assertion commands.
We write $\setssmp$ for the set of simple sequences.
\end{mydef}

>>>>

% TODO Define, spell out update decetion algorithm

\begin{mydef}[Minimal sets or sequences]
A sequence or set of commands is minimal if it contains at most one command on each node.
\end{mydef}

\begin{mydef}[Simple sets or sequences]
A sequence or set of commands is simple if it is minimal and it does not contain assertion commands.
\end{mydef}
